---
title: "bmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(drc)
library(bmd)
library(proast70.0)
library(assertive)
library(hwriter)
install.packages(c('assertive', 'gWidgetstcltk', 'hwriter'))
```


Developmental dose-response data on cleft palate for two compounds believed to operate via similar
mechanisms of action.


The data are part of a larger collection of datasets reported in Sand et al. (2002).

Sand, S., Falk Fillipsson, A. and Victorin, K. (2002) Evaluation of the Benchmark Dose Method for
Dichotomous Data: Model Dependence and Model Selection, Regulatory Toxicology and Pharmacology, 36, 184–197

```{r }

cleft.palate


tcdd <-
  as_tibble(cleft.palate) %>% 
  mutate(response = affected / total)


# bar plot
tcdd %>% 
  ggplot(aes(x = dose, y = response)) +
    geom_bar(stat = "identity") +
    facet_wrap(~compound) +
    theme_bw()

# data points
tcdd %>% 
  ggplot(aes(x = dose, y = response)) +
    geom_point(size = 4) +
    facet_wrap(~compound) +
    theme_bw()  


?drm

```

Fit model to TCDD1

```{r}

tcdd1 <-
  tcdd %>% 
    filter(compound %in% "2,3,7,8-TCDD1")


tcdd2 <-
  tcdd %>% 
    filter(compound %in% "2,3,7,8-TCDD2")



## Fitting log-logistic two-parameter model to binomial data

dr_tcdd1_m1 <-
  drc::drm(response ~ dose, 
           weights = total,
           data = tcdd1,
           fct = LL.2(),
           type = "binomial")

dr_tcdd1_m1
summary(dr_tcdd1_m1)


plot(dr_tcdd_m1)

## Estimated TD values
ED(dr_tcdd1_m1, c(5, 10, 50), interval = "delta")


## Excess risk with BMR=5% and background 0

bmd(dr_tcdd1_m1, c(0.05, 0.1))

```


PROAST
The National Institute for Public Health and the Environment of the Netherlands
(RIVM) software for dose-response modelling and benchmark dose analysis (PROAST) 

```{r}

data(das11)
das11


f.proast(das11)

# 1,3,1,13(relative kidney weights), 1, 0, 2, 200, 2

```



