---
title: "monte_carlo"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(mc2d)
```

## Arsenic Risk Example - Monte Carlo Simulation

Define variables and distribution parameters

Term	Abv	Units	Fixed/Variable	Input Value	Distribution	Param 1	Param 2	Param 3
Inputs								
Concentration	C	mg/L	variable	0.00375	Lognormal	0.00375	0.001	
Intake Rate	IR	L/day	variable	1	Uniform	0.5	5	
Exposure Frequency	EF	days/yr	variable	350	Beta-PERT	200	350	365
Exposure Duration	ED	years	fixed	6				
Body Weight	BW	kg	variable	70	Normal	70	25	
Averaging Time	AT	days	fixed	25550				
								
Cancer Slope Factor	CSF	1 / mg/kg-d	fixed	1.5				


```{r}
c <- 0.00375
c_p1 <- 0.00375
c_p2 <- 0.001
ir <- 1
ir_p1 <- 0.5
ir_p2 <- 1
ir_p3 <- 5
ef <- 350
ef_p1 <- 200
ef_p2 <- 350
ef_p3 <- 365
ed <- 6
bw <- 70
bw_p1 <- 70
bw_p2 <- 75
at <- 365*70
csf <- 1.5

```

## Deterministic
Daily Intake = (C x IR x EF x ED x) / (BW x AT)
Risk = CSF x (C x IR x EF x ED x) / (BW x AT)
```{r}
daily_intake <-
  (c * ir * ef * ed) / (bw * at) 

daily_intake

risk <-
  csf * ((c * ir * ef * ed) / (bw * at)) 

risk

```

## Probabilistic

```{r}


# Concentration (C): Lognormal

# reparameterize for `rlnorm`
c_location <- log(c_p1^2 / sqrt(c_p2^2 + c_p1^2))
c_shape <- sqrt(log(1 + (c_p2^2 / c_p1^2)))

c_d <- rlnorm(10000, c_location, c_shape)
# c %>% as_tibble()
hist(c)


# Intake Rate (IR): Beta-PERT
ir_d <- mc2d::rpert(n = 10000, min = ir_p1, mode = ir_p2, max = ir_p3, shape = 4)
  
# ir %>% as_tibble()
hist(ir)

```


Replicate and Solve for Risk

```{r}

risk_mc <-
  replicate(n = 10000, expr = {
    c_i = sample(c_d, size = 1)
    ir_i = sample(ir_d, size = 1)
    csf * ((c_i * ir_i * ef * ed) / (bw * at)) 
  })

risk_mc %>%
  as_tibble() %>% 
  ggplot(aes(x = value)) +
  geom_histogram(bins = 60) +
  theme_bw() +
  xlab("risk")


risk_mc %>%
  as_tibble() %>% 
  ggplot(aes(x = value)) +
  geom_boxplot() +
  theme_bw() +
  xlab("risk")

summary(risk_mc)
```


